Bayesian Inference in Numerical Cognition: A Tutorial Using JASP

### 2020-03-08 jncTutorial/manuscript.Rmd
add submitted version of tutorial

abstract: >
  Researchers in numerical cognition rely on hypothesis testing and estimation to evaluate the evidential value of data. Though there has been increased interest in Bayesian statistics as an alternative to the classical, frequentist approach to hypothesis testing, many researchers remain hesitant to change their methods of inference. In this tutorial, we provide a concise introduction to Bayesian hypothesis testing and estimation in the context of numerical cognition. Here, we focus on three examples of Bayesian inference: the \( t \)-test, linear regression, and analysis of variance. Using the free software package JASP, we provide the reader with a basic understanding of how Bayesian inference works "under the hood" as well as instructions detailing how to perform and interpret each Bayesian analysis.

bibliography      : ["references.bib"]

@@

lang              : "english"
class             : "doc" # can change this to "man" or "jou" or "doc"
output            : papaja::apa6_pdf
---

@@

Like any other discipline within the psychological and behavioral sciences, researchers in numerical cognition rely on statistical inference, and hypothesis testing in particular, to verify theoretical claims against observed data. In recent years, Bayesian inference has become increasingly popular as an alternative to classical null hypothesis significance testing (NHST). Bayesian inference operates under a fundamentally different framework from NHST; instead of hard "accept/reject" decisions about a null hypothesis, Bayesian inference works by quantifying evidence about competing hypotheses in light of observed data. Hypotheses that do a better job of predicting data will receive increased support, whereas hypotheses that do not predict the data well will receive decreased support.

@@

## Hypotheses
Hypothesis testing can be viewed as a process of validating general claims based on a small sample of the general population. To illustrate, suppose that we are policy makers confronted with a new study programme which claims to increase the mathematical abilities of sixth graders in the United States. To ease the exposition we use a (fictitious) scale called the Scale for Advancing Mathematical Ability (SAMA) to measure mathematical ability. As policy makers we know that the population mean SAMA score is \( \mu_{0}=50 \) and that the standard deviation is \( \sigma=10 \) for the population of sixth graders that are taught using the standard materials in the US.

The superiority claim about the new study programme can be formalized as the working hypothesis \( \mathcal{H}_{1} : \mu > 50 \). Here \( \mu \) refers to the \emph{population} mean of SAMA score of the same sixth graders, but taught using the new study program instead. Such a treatment population does not (yet) exist and \( \mu \) cannot be measured, and is, therefore, unknown.

Nonetheless, we would like to infer the validity of the working hypothesis \( \mathcal{H}_{1} \) by running an experiment in which, say, \( n=90 \) sixth graders are taught with the new study programme. Suppose that these particular participants scored an average SAMA score of \( \bar{y} = 52.07 \). Clearly, this sample mean is larger than the status quo population mean of \( 50 \), but can we assume that this finding generalizes to the whole population? If the answer is "yes" and we declare the working hypothesis to be verified, then how confident would we be about such a decision? Below we recall how \( p \)-values are typically used to make hard "yes" and "no" decisions, and we remind the reader why a \( p \)-value does not quantify the (un)certainty of such decisions.

## Statistical models
To deduce how hypotheses, i.e., claims about the population such as \( \mu \), affect the data, such as \( \bar{y} \), we have to link the former to the latter with a distribution. For instance, let us suppose a normal distribution provides this link. The population parameters, the distribution that links them to the data, and the data together make up a (statistical) model, which allows for quantitative predictions. For instance, the claim that the treatment population is identitical to the current population of sixth graders implies the null model \( \mathcal{M}_{0} \) comprised of (1) the null hypothesis \( \mathcal{H}_{0}: \mu = 50 \), (2) the assumption \( \sigma = 10 \), and (3) normally distributed SAMA scores. When all three assumptions hold true, we can mathematically deduce that the \emph{potential} average SAMA score for a group of \( n = 90 \) participants is expected to fall within the interval \( 47.93 = 50 - 1.96 \frac{10}{\sqrt{n}} \) and \( 52.07 = 50 + 1.96 \frac{10}{\sqrt{n}} \) with 95% chance.<!--
-->
\footnote{We use ``chance'' to refer to probabilistic statements about the data.}
<!--
-->

@@

Under the null model \( \mathcal{M}_{0} \), the more extreme data are potential outcomes of \( \bar{y} \) larger than the observed sample mean of \( 52.07 \), but also lower than the unobserved possible sample mean of \( 47.93 = 50 - (52.07-50) \) (i.e., a potential sample mean that is less than, but equally far from, the hypothesized mean of 50) . Under the null, and the normality assumption in particular, we can deduce that the chance of seeing a potential \( \bar{y} \) in \( (- \infty, 47.93] \) or \( [52.07, \infty) \) is 5%, thus, \( p=0.05 \). Note that \( 1 - p = 0.95 \), which is the 95% chance of the non-extreme event of having \( \bar{y} \) fall within the interval \( [47.93, 52.07] \), if the null model holds true. We use this example to emphasize that \( p \)-values help us reason from the null model to (potential) data, but they do not allow us to do the reverse; that is, make quantitative statements from the actually observed data about the hypotheses.

A small \( p \)-value has many potential causes, and for this case we mention four: (1) the null hypothesis is not true; (2) the true variance is actually higher than \( 10 \); (3) the data are not normally distributed; or (4) the null model actually holds true, but we saw a rare event. Despite the various causes of a small \( p \), typically only (1) is singled out and mentioned as \emph{the} culprit. The typical conclusion is then to reject the null, and the working hypothesis \( \mathcal{H}_{1} :  \mu > 50 \) is proposed as the winner. Note that this type of reasoning provides only an indirect argument for our working hypothesis, as the predictions of the working hypothesis itself were not considered at all. Only the null was used in the computation of \( p \). The fact that the null is not a good fit for the data, does not automatically imply that the alternative fit the data well.

We conclude our brief review on \( p \)-values with three observations: (i) \( p \)-values only allow us to deduce predictions from the general population to potential data, but not the other way around; (ii) the comparison between \( \mathcal{H}_{0} \) and \( \mathcal{H}_{1} \) is unfair in the sense that \( p \)-values only consider one side of the coin (namely, the fit of \( \mathcal{H}_{0} \)), and thus, support for \( \mathcal{H}_1 \) is only ever indirect; and (iii) traditional practice\footnote{From the Fisherian point of view, $p$-values are considered as a continuous measure of evidence against the null. However, in the Neyman-Pearson testing framework, the null is accepted or rejected based on whether the test statistic reaches some critical value. In practice, researchers have mixed these paradigms, leading to the common practice of using $p$-values as a basis for making accept/reject decisions. } with \( p \)-values can only lead to hard reject or non-reject decisions, which encourages "black and white" thinking and ignores the nuances of a scientific enquiry. On this issue, @rozeboom1960 writes

\begin{quote}
The null-hypothesis significance test treats "acceptance" or "rejection" of a hypothesis as though these were decisions one makes. But a hypothesis is not something, like a piece of pie offered for dessert, which can be accepted or rejected by a voluntary physical action. Acceptance or rejection of a hypothesis is a cognitive process, a degree of believing or disbelieving which, if rational, is not a matter of choice but determined solely by how likely it is, given the evidence, that the hypothesis is true (p. 422-423).
\end{quote}

@@

\begin{quote}
Modern statisticians have developed extensive mathematical techniques, but for the most part have rejected the notion of the probability of a hypothesis, and thereby deprived themselves of any way of saying precisely what they mean when they decide between hypotheses (p. v).
\end{quote}

Though there are many other problems with $p$-values [see @wagenmakers2007], we contend that the key point is the neglect of any indication of support for $\mathcal{H}_1$. One way to remedy this is to consider Bayesian hypothesis testing as an alternative.

## Bayesian evidence as relative predictive performance
To address these problems, we can use Bayesian hypothesis testing, which gives us the ability to quantify evidence by comparing the *relative predictive performance* of both $\mathcal{H}_0$ and $\mathcal{H}_1$. Compared to null-hypothesis testing, Bayesian testing adds the missing piece $p(\text{data}\mid \mathcal{H}_1)$, which indexes the predictive performance of $\mathcal{H}_1$. We then quantify the relative predictive performance by computing the *Bayes factor* [@kass1995],
\[
\text{BF}_{10} = \frac{p(\text{data}\mid\mathcal{H}_1)}{p(\text{data}\mid\mathcal{H}_0)},
\]
which indexes the relative adequacy of both hypotheses as predictors of our observed data. One immediate benefit of the Bayes factor is ease of interpretation. For example, if $\text{BF}_{10}=5$, this tells us that our data are 5 times more likely to be observed under $\mathcal{H}_1$ than $\mathcal{H}_0$. Thus, we go beyond $p$-values by simultaneously assessing the predictive adequacy of *both* hypotheses, not just $\mathcal{H}_0$.  Of course, in order for $\mathcal{H}_1$ to make predictions, we have to be specific about what is meant by "$\mu>50$". Because we are rarely certain about what "$\mu>50$" means exactly, we place a *prior* distribution on the parameter $\mu$. We will specify more about this in later when we get to specific tests, but in general, the prior allows us to express our uncertainty about the values of $\mu$ under $\mathcal{H}_1$. For now, please note that this prior is a technical piece that is necessary for computing the Bayes factor.

@@

Unlike the \( p \)-value, there are no universal standards for when a Bayes factor provides a sufficient amount of evidence. However, people have made recommendations in the past regarding sizes of Bayes factors, and these provide us with a helpful, but very rough guideline on its interpretation, when we do not know the context of the problem. In particular, we mention the recommendations of @jeffreys1961, who proposed the following scheme<!--
-->
\footnote{We advise readers not to get too hung up on the specific labels and breakpoints in this table, see also @giollaInpresswhat .}:

\begin{center}
\begin{tabular}{rl}
Bayes factor & Evidence\\
\hline
1-3 & anecdotal\\
3-10 & moderate\\
10-30 & strong\\
30-100 & very strong\\
\( > \)100 & extreme\\
\end{tabular}
\end{center}

In principle, the Bayes factor does not come with a specified decision rule. We believe that context made explicit using prior model probabilities by field experts are much more important, than overly generalized "rules" such \( p < 0.05 \) or \( \text{BF} _{10} > k \), where \( k \) is some arbitrarily chosen number.

It is also worth nothing that a Bayes factor is simply a quantity which measures relative evidence between two models. It is completely independent of our prior beliefs in the two hypotheses $\mathcal{H}_0$ and $\mathcal{H}_1$. To incorporate the Bayes factor (relative predictive performance) with these beliefs (i.e., the prior model probabilities), we need Bayes' rule.

@@

Bayes' rule mathematically expresses the most basic idea in science; namely, that our prior belief in a hypothesis, $p(\mathcal{H})$, is updated after observing data to a *posterior* belief $p(\mathcal{H}\mid \text{data})$ in the following way:

\[
\underbrace{P(\mathcal{H}\mid \text{data})}_{\substack{\text{Posterior belief}\\ \text{about hypothesis}}} = \underbrace{P(\mathcal{H})}_{\substack{\text{Prior belief}\\ \text{about hypothesis}}} \times \underbrace{\frac{P(\text{data}\mid \mathcal{H})}{P(\text{data})}}_{\text{predictive updating factor}}.
\]

Since our goal is to directly compare the predictive adequacy of two hypotheses $\mathcal{H}_0$ and $\mathcal{H}_1$, we can employ Bayes' rule twice to derive the following:

\begin{align*}
  \frac{P(\mathcal{H}_1\mid \text{data})}{P(\mathcal{H}_0\mid \text{data})}
  & = \frac{P(\mathcal{H}_1) \cdot \frac{P(\text{data}\mid \mathcal{H}_1)}{P(\text{data})}}{P(\mathcal{H}_0) \cdot  \frac{P(\text{data}\mid \mathcal{H}_0)}{P(\text{data})}} \\ 
  & = \frac{P(\mathcal{H}_1)\cdot P(\text{data}\mid \mathcal{H}_1)}{P(\mathcal{H}_0)\cdot P(\text{data}\mid \mathcal{H}_0)},\\
\end{align*}

which we can simplify to

\[
\underbrace{\frac{P(\mathcal{H}_1\mid \text{data})}{P(\mathcal{H}_0\mid \text{data})}}_{\substack{\text{posterior beliefs}\\ \text{about hypotheses}}} = \underbrace{\frac{P(\mathcal{H}_1)}{P(\mathcal{H}_0)}}_{\substack{\text{prior beliefs}\\ \text{about hypotheses}}} \times \underbrace{\frac{P(\text{data}\mid \mathcal{H}_1)}{P(\text{data}\mid \mathcal{H}_0)}}_{\text{predictive updating factor}}, 
\]

or equivalently,

\[
\frac{P(\mathcal{H}_1\mid \text{data})}{P(\mathcal{H}_0) \mid \text{data})} = \frac{P(\mathcal{H}_1)}{P(\mathcal{H}_0)} \times \text{BF}_{10}. 
\]

This gives us a lovely way to remember another fundamental fact of Bayesian hypothesis testing; that is, *posterior odds = prior odds $\times$ BF*. Thus, not only does the Bayes factor assess the relative predictive performance of two hypotheses, it also indexes the factor by which our relative belief in the two hypotheses can be updated after observing data.

Now we come back to our earlier example of the SAMA test for mathematical ability and demonstrate these ideas in context. Working as policy makers, we might have seen many claims of improved study programs that did not work out well. For instance, if past claims of improved study programmes led to null results four times out of five, we might place 4-to-1 odds against the claim of improvement and thus set \( P( \mathcal{H}_{1}) = 0.2 \) and \( P(\mathcal{H}_{0}) = 0.8 \). On the other hand, the designers of the new study programme to increase SAMA scores might feel strongly about the benefits of their programme, and so might choose \( P( \mathcal{H}_{1}) = 0.7 \) and \( P(\mathcal{H}_{0}) = 0.3 \) instead. If \( \mathcal{H}_{1} \) and \( \mathcal{H}_{0} \) are the only models under consideration we have \( 0 < P( \mathcal{H}_{0}) < 1 \) and \( P( \mathcal{H}_{1})= 1 - P( \mathcal{H}_{0}) \). The Bayes factor \( \text{BF}_{10}(d) \) can then be used to update these prior model probabilities to posterior model probabilities, as follows:

\begin{equation}
\label{eqPriorToPosterior}
P(\mathcal{H}_{1} \mid \text{data}) = \frac{ \text{BF}_{10} P(\mathcal{H}_{1})}{\text{BF}_{10} P(\mathcal{H}_{1}) + P(\mathcal{H}_{0})} \text{, and } P(\mathcal{H}_{0} \mid \text{data}) = 1 - P(\mathcal{H}_{1} \mid \text{data}).
\end{equation}

For illustration, suppose that we found \( \text{BF}_{10} = 5 \). Then the sceptic (who set $P(\mathcal{H}_1=0.2$ before seeing the data) increases their belief in the superiority of the new study programme to 56%, i.e., \( P(\mathcal{H}_{1} \mid \text{data})=0.56 \), and retains 44% probability of for the null after data observation. On the other hand, the same \( \text{BF}_{10} = 5 \) results in the proponent of the new study programme (who set $P(\mathcal{H}_1)=0.7$ beofre seeing the data) to report a posterior model probability of \( P(\mathcal{H}_{1} \mid \text{data})=0.92 \) and \( P(\mathcal{H}_{0} \mid \text{data})=0.08 \) instead. Note that the posterior model probabilities combine what we believed prior to data observation with the evidence of the data in terms of the Bayes factors. Both sceptic and the proponent could, if they wish, declare the new study programme to be superior to the status quo, but the sceptic would be much less confident than the proponent. The resulting posterior model probabilities can, thus, be used to complement a decision with a measure of uncertainty.

@@

As a general method, Bayesian hypothesis testing gives us several pragmatic benefits, each of which will be highlighted in our examples below. The first of these is the ability to quantify evidence on a continuous scale.<!--For a given set of observed data, the Bayes factor indexes the relative predictive adequacy of $\mathcal{H}_0$ and $\mathcal{H}_1$. For example, if $\text{BF}_{10} > 1$, the data are more likely under $\mathcal{H}_1$ than they are under $\mathcal{H}_0$, and are thus evidential for $\mathcal{H}_1$. Further, the value of the Bayes factor indexes the amount of this evidence on a continuous scale.--> A second benefit is the ability to differentiate between *evidence of absence* and *absence of evidence*. Recall that in traditional null hypothesis testing, failure to reject $\mathcal{H}_0$ cannot be taken as evidence for a null effect. In this instance, it is not clear whether our data provide evidence for $\mathcal{H}_0$ (i.e., evidence of absence) or our data are simply not evidential for either hypothesis (i.e., absence of evidence). The Bayes factor gives us the ability to separate these two ideas. For example, if $\text{BF}_{01}$ is large, we take this to represent evidence for the null -- that is, evidence for the absence of something.  Another benefit include the ability for the researcher to monitor evidence as the data accumulate. Though problematic for inference based on $p$-values, optional stopping does not affect the interpretation of Bayesian analyses [@rouder2014optional]. Thus, researchers may continuously monitor the "evidential flow" of their data and choose to stop whenever evidence is sufficiently compelling. Finally, Bayesian methods integrate testing with estimation, allowing the researcher to not only measure the evidence for one hypothesis over another, but to simultaneously obtain a probabilistic estimate of the effect size under the hypothesis that it is non-zero.

Given these benefits, what does a Bayesian hypothesis test require of the researcher? Like any hypothesis testing situation, the researcher must specify a null hypothesis $\mathcal{H}_0$ and an alternative hypothesis $\mathcal{H}_1$. In a Bayesian context, though, there are some specific considerations that govern how this choice is made. Essentially, the researcher must specify a *prior distribution* of effect sizes that are expected under $\mathcal{H}_1$. This allows $\mathcal{H}_1$ to makes predictions that can be evaluated against observed data. For example, the Bayesian $t$-test of @rouder2009bayesian uses a Cauchy distribution with scale $r=1/\sqrt{2}$ as a default choice\footnote{This prior satisfies several desirable properties for an objective, default prior; see the supplemental material for details.} for the prior distribution of effect sizes that are expected *if* there is a nonzero effect. Given this definition of $\mathcal{H}_1$ (i.e., as a distribution of effect sizes $\delta$), the null hypothesis $\mathcal{H}_0$ is specified as a specific point within this distribution, namely $\delta=0$. Thus, we can succinctly encode our hypotheses as:

>>>>

abstract: >
  Researchers in numerical cognition rely on hypothesis testing and parameter estimation to evaluate the evidential value of data. Though there has been increased interest in Bayesian statistics as an alternative to the classical, frequentist approach to hypothesis testing, many researchers remain hesitant to change their methods of inference. In this tutorial, we provide a concise introduction to Bayesian hypothesis testing and parameter estimation in the context of numerical cognition. Here, we focus on three examples of Bayesian inference: the \( t \)-test, linear regression, and analysis of variance. Using the free software package JASP, we provide the reader with a basic understanding of how Bayesian inference works "under the hood" as well as instructions detailing how to perform and interpret each Bayesian analysis.

bibliography      : ["referencesShort.bib"]

@@

lang              : "english"
class             : "man" # can change this to "man" or "jou" or "doc"
output            : papaja::apa6_pdf
---

@@

Like any other discipline within the psychological and behavioral sciences, researchers in numerical cognition rely on statistical inference -- hypothesis testing and parameter estimation -- to verify theoretical claims against observed data. In recent years, Bayesian inference has become increasingly popular as an alternative to classical null hypothesis significance testing (NHST). Bayesian inference operates under a fundamentally different framework from NHST; instead of hard "accept/reject" decisions about a null hypothesis, Bayesian inference works by quantifying evidence about competing hypotheses in light of observed data. Hypotheses that do a better job of predicting data will receive increased support, whereas hypotheses that do not predict the data well will receive decreased support.

@@

## Hypotheses
Hypothesis testing can be viewed as a process of validating general claims based on a small sample of the general population. To illustrate, suppose that we are policy makers confronted with a new study program which claims to increase the mathematical abilities of sixth graders in the United States. To ease the exposition we use a (fictitious) scale called the Scale for Advancing Mathematical Ability (SAMA) to measure mathematical ability. As policy makers we know that the population mean SAMA score is \( \mu_{0}=50 \) and that the standard deviation is \( \sigma=10 \) for the population of sixth graders that are taught using the standard materials in the US.

The superiority claim about the new study program can be formalized as the working hypothesis \( \mathcal{H}_{1}:\mu > 50 \). Here \( \mu \) refers to the *population* mean of SAMA score of the same sixth graders, but taught using the new study program instead. Such a treatment population does not (yet) exist and \( \mu \) cannot be measured, and is, therefore, unknown.

Nonetheless, we would like to infer the validity of the working hypothesis \( \mathcal{H}_{1} \) by running an experiment in which, say, \( n=90 \) sixth graders are taught with the new study program. Suppose that these particular participants scored an average SAMA score of \( \bar{y} = 52.07 \). Clearly, this sample mean is larger than the status quo population mean of \( 50 \), but can we assume that this finding generalizes to the whole population? If the answer is "yes" and we declare the working hypothesis to be verified, then how confident would we be about such a decision? Below we recall how \( p \)-values are typically used to make hard "yes" and "no" decisions, and we remind the reader why a \( p \)-value does not quantify the (un)certainty of such decisions.

## Statistical models
To deduce how hypotheses, i.e., claims about the population such as \( \mu \), affect the data, such as \( \bar{y} \), we have to link the former to the latter with a distribution. For instance, let us suppose a normal distribution provides this link. The population parameters, the distribution that links them to the data, and the data together make up a (statistical) model, which allows for quantitative predictions. For instance, the claim that the treatment population is identitical to the current population of sixth graders implies the null model \( \mathcal{M}_{0} \) comprised of (1) the null hypothesis \( \mathcal{H}_{0}: \mu = 50 \), (2) the assumption \( \sigma = 10 \), and (3) normally distributed SAMA scores. When all three assumptions hold true, we can mathematically deduce that the \emph{potential} average SAMA score for a group of \( n = 90 \) participants is expected to fall within the interval \( 47.93 = 50 - 1.96 \frac{10}{\sqrt{n}} \) and \( 52.07 = 50 + 1.96 \frac{10}{\sqrt{n}} \) with 95% chance.^[We use ``chance'' to refer to probabilistic statements about the data.]

@@

Under the null model \( \mathcal{M}_{0} \), the more extreme data consist of potential outcomes of \( \bar{y} \) larger than the observed sample mean of \( 52.07 \), as well as those potential values of \( \bar{y} \) that are smaller than the unobserved possible sample mean of \( 47.93 = 50 - (52.07-50) \) (i.e., a potential sample mean that is less than, but equally far from, the hypothesized mean of 50) . Under the null, and the normality assumption in particular, we can deduce that the chance of seeing a potential \( \bar{y} \) in \( (- \infty, 47.93] \) or \( [52.07, \infty) \) is 5%, thus, \( p=0.05 \). Note that \( 1 - p = 0.95 \), which is the 95% chance of the non-extreme event of having \( \bar{y} \) fall within the interval \( [47.93, 52.07] \), if the null model holds true. We use this example to emphasize that \( p \)-values help us reason from the null model to (potential) data, but they do not allow us to do the reverse; that is, make quantitative statements from the actually observed data about the hypotheses.

A small \( p \)-value has many potential causes, and for this case we mention four: (1) the null hypothesis is not true; (2) the true variance is actually higher than \( 10 \); (3) the data are not normally distributed; or (4) the null model actually holds true, but we saw a rare event. Despite the various causes of a small \( p \), typically only (1) is singled out and mentioned as \emph{the} culprit. The typical conclusion is then to reject the null, and the working hypothesis \( \mathcal{H}_{1} :  \mu > 50 \) is proposed as the winner. Note that this type of reasoning provides only an *indirect* argument for our working hypothesis, as the predictions of the working hypothesis itself were not considered at all. Only the null was used in the computation of \( p \). The fact that the null is not a good fit for the data does not automatically imply that the alternative fit the data well.

We conclude our brief review on \( p \)-values with three observations: (i) \( p \)-values only allow us to deduce predictions from the general population to potential data, but not the other way around; (ii) the comparison between \( \mathcal{H}_{0} \) and \( \mathcal{H}_{1} \) is unfair in the sense that \( p \)-values only consider one side of the coin -- namely, the fit of \( \mathcal{H}_{0} \) -- and thus, support for \( \mathcal{H}_1 \) is only ever indirect; and (iii) traditional practice^[From the Fisherian point of view, $p$-values are considered as a continuous measure of evidence against the null. However, in the Neyman-Pearson testing framework, the null is accepted or rejected based on whether the test statistic reaches some critical value. In practice, researchers have mixed these paradigms, leading to the common practice of using $p$-values as a basis for making accept/reject decisions.] with \( p \)-values can only lead to hard reject or non-reject decisions, which encourages "black and white" thinking and ignores the nuances of a scientific enquiry. On this issue, @rozeboom1960 writes

> The null-hypothesis significance test treats "acceptance" or "rejection" of a hypothesis as though these were decisions one makes. But a hypothesis is not something, like a piece of pie offered for dessert, which can be accepted or rejected by a voluntary physical action. Acceptance or rejection of a hypothesis is a cognitive process, a degree of believing or disbelieving which, if rational, is not a matter of choice but determined solely by how likely it is, given the evidence, that the hypothesis is true (p. 422-423).

@@

> Modern statisticians have developed extensive mathematical techniques, but for the most part have rejected the notion of the probability of a hypothesis, and thereby deprived themselves of any way of saying precisely what they mean when they decide between hypotheses (p. v).

Though there are many other problems with \( p \)-values [see @wagenmakers2007], we contend that the key point is the neglect of any indication of support for \( \mathcal{H}_1 \). One way to remedy this is to consider Bayesian hypothesis testing as an alternative.

## Bayesian evidence as relative predictive performance
To address these problems, we can use Bayesian hypothesis testing, which gives us the ability to quantify evidence by comparing the *relative predictive performance* of both \( \mathcal{H}_0 \) and \( \mathcal{H}_1 \). Compared to null-hypothesis testing, Bayesian testing adds the missing piece \( p(\text{data}\mid \mathcal{H}_1) \), which indexes the predictive performance of \( \mathcal{H}_1 \). We then quantify the relative predictive performance by computing the *Bayes factor* [@kass1995],<!--
-->
\[
\text{BF}_{10} = \frac{p(\text{data}\mid\mathcal{H}_1)}{p(\text{data}\mid\mathcal{H}_0)},
\]<!--
-->
which indexes the relative adequacy of both hypotheses as predictors of our observed data. One immediate benefit of the Bayes factor is ease of interpretation. For example, if \( \text{BF}_{10}=5 \), this tells us that our data are 5 times more likely to be observed under \( \mathcal{H}_1 \) than \( \mathcal{H}_0 \). Thus, we go beyond \( p \)-values by simultaneously assessing the predictive adequacy of *both* hypotheses, not just \( \mathcal{H}_0 \).  Of course, in order for \( \mathcal{H}_1 \) to make predictions, we have to be specific about what is meant by "\( \mu > 50 \)". Because we are rarely certain about what "\( \mu > 50 \)" means exactly, we place a *prior* distribution on the parameter \( \mu \). We will specify more about this in later when we get to specific tests, but in general, the prior allows us to mathematically express our uncertainty about the values of \( \mu \) under \( \mathcal{H}_1 \). For now, please note that this prior is a technical piece that is necessary for computing the Bayes factor.

@@

Unlike the \( p \)-value, there are no universal standards for when a Bayes factor provides a sufficient amount of evidence. However, people have made recommendations in the past regarding sizes of Bayes factors, and these provide us with a helpful, but very rough guideline on its interpretation, when we do not know the context of the problem. In particular, we mention the recommendations of @jeffreys1961, who proposed the following scheme^[We advise readers not to get too hung up on the specific labels and breakpoints in this table, see also @giollaInpresswhat]:<!--
-->
\begin{center}
\begin{tabular}{rl}
Bayes factor & Evidence\\
\hline
1-3 & anecdotal\\
3-10 & moderate\\
10-30 & strong\\
30-100 & very strong\\
\( > \)100 & extreme\\
\end{tabular}
\end{center}<!--
-->
In principle, the Bayes factor does not come with a specified decision rule. We believe that context made explicit using prior model probabilities by field experts are much more important than overly generalized "rules" such \( p < 0.05 \) or \( \text{BF} _{10} > k \), where \( k \) is some arbitrarily chosen number. It is also worth nothing that a Bayes factor is simply a quantity which measures relative evidence between two models. It is completely independent of our prior beliefs in the two hypotheses \( \mathcal{H}_0 \) and \( \mathcal{H}_1 \). To incorporate the Bayes factor (relative predictive performance) with these beliefs (i.e., the prior model probabilities), we need Bayes' rule.

@@

Bayes' rule mathematically expresses the most basic idea in science; namely, that our prior belief in a hypothesis -- \( p(\mathcal{H}) \) -- is updated after observing data to a *posterior* belief \( p(\mathcal{H}\mid \text{data}) \) in the following way:<!--
-->
\[
\underbrace{P(\mathcal{H}\mid \text{data})}_{\substack{\text{Posterior belief}\\ \text{about hypothesis}}} = \underbrace{P(\mathcal{H})}_{\substack{\text{Prior belief}\\ \text{about hypothesis}}} \times \underbrace{\frac{P(\text{data}\mid \mathcal{H})}{P(\text{data})}}_{\text{predictive updating factor}}.
\]<!--
-->
Since our goal is to directly compare the predictive adequacy of two hypotheses \( \mathcal{H}_0 \) and \( \mathcal{H}_1 \), we can employ Bayes' rule twice to derive the following:<!--
-->
\begin{align*}
  \frac{P(\mathcal{H}_1\mid \text{data})}{P(\mathcal{H}_0\mid \text{data})}
  & = \frac{P(\mathcal{H}_1) \cdot \frac{P(\text{data}\mid \mathcal{H}_1)}{P(\text{data})}}{P(\mathcal{H}_0) \cdot  \frac{P(\text{data}\mid \mathcal{H}_0)}{P(\text{data})}} \\ 
  & = \frac{P(\mathcal{H}_1)\cdot P(\text{data}\mid \mathcal{H}_1)}{P(\mathcal{H}_0)\cdot P(\text{data}\mid \mathcal{H}_0)},\\
\end{align*}<!--
-->
which we can simplify to<!--
-->
\[
\underbrace{\frac{P(\mathcal{H}_1\mid \text{data})}{P(\mathcal{H}_0\mid \text{data})}}_{\substack{\text{posterior beliefs}\\ \text{about hypotheses}}} = \underbrace{\frac{P(\mathcal{H}_1)}{P(\mathcal{H}_0)}}_{\substack{\text{prior beliefs}\\ \text{about hypotheses}}} \times \underbrace{\frac{P(\text{data}\mid \mathcal{H}_1)}{P(\text{data}\mid \mathcal{H}_0)}}_{\text{predictive updating factor}}, 
\]<!--
-->
or equivalently,<!--
-->
\[
\frac{P(\mathcal{H}_1\mid \text{data})}{P(\mathcal{H}_0) \mid \text{data})} = \frac{P(\mathcal{H}_1)}{P(\mathcal{H}_0)} \times \text{BF}_{10}. 
\]<!--
-->
This gives us a lovely way to remember another fundamental fact of Bayesian hypothesis testing; that is, *posterior odds = prior odds $\times$ BF*. Thus, not only does the Bayes factor assess the relative predictive performance of two hypotheses, it also indexes the factor by which our relative belief in the two hypotheses can be updated after observing data.

Now we come back to our earlier example of the SAMA test for mathematical ability and demonstrate these ideas in context. Working as policy makers, we might have seen many claims of improved study programs that did not work out well. For instance, if past claims of improved study programs led to null results four times out of five, we might place 4-to-1 odds *against* the claim of improvement and thus set \( P( \mathcal{H}_{1}) = 0.2 \) and \( P(\mathcal{H}_{0}) = 0.8 \). On the other hand, the designers of the new study program to increase SAMA scores might feel strongly about the benefits of their program, and so instead might choose \( P( \mathcal{H}_{1}) = 0.7 \) and \( P(\mathcal{H}_{0}) = 0.3 \). If \( \mathcal{H}_{1} \) and \( \mathcal{H}_{0} \) are the only models under consideration we have \( 0 < P( \mathcal{H}_{0}) < 1 \) and \( P( \mathcal{H}_{1})= 1 - P( \mathcal{H}_{0}) \). The Bayes factor \( \text{BF}_{10} \) can then be used to update these prior model probabilities to posterior model probabilities, as follows:<!--
-->
\[
P(\mathcal{H}_{1} \mid \text{data}) = \frac{ \text{BF}_{10} P(\mathcal{H}_{1})}{\text{BF}_{10} P(\mathcal{H}_{1}) + P(\mathcal{H}_{0})} \text{, and } P(\mathcal{H}_{0} \mid \text{data}) = 1 - P(\mathcal{H}_{1} \mid \text{data}).
\]<!--
-->
For illustration, suppose that we found \( \text{BF}_{10} = 5 \). Then the skeptic (who set \( P(\mathcal{H}_1=0.2 \) before seeing the data) increases their belief in the superiority of the new study program to 56%, i.e., \( P(\mathcal{H}_{1} \mid \text{data})=0.56 \), and retains 44% probability of for the null after data observation. On the other hand, the same \( \text{BF}_{10} = 5 \) results in the proponent of the new study program (who set \( P(\mathcal{H}_1)=0.7 \) before seeing the data) to report a posterior model probability of \( P(\mathcal{H}_{1} \mid \text{data})=0.92 \) and \( P(\mathcal{H}_{0} \mid \text{data})=0.08 \) instead. Note that the posterior model probabilities combine what we believed prior to data observation with the evidence of the data in terms of the Bayes factors. Both the skeptic and the proponent could, if they wish, declare the new study program to be superior to the status quo, but the skeptic would be much less confident than the proponent. The resulting posterior model probabilities can, thus, be used to complement a decision with a measure of uncertainty.

@@

As a general method, Bayesian hypothesis testing gives us several pragmatic benefits, each of which will be highlighted in our examples below. The first of these is the ability to quantify evidence on a continuous scale. A second benefit is the ability to differentiate between *evidence of absence* and *absence of evidence*. Recall that in traditional null hypothesis testing, failure to reject \( \mathcal{H}_0 \) cannot be taken as evidence for a null effect. In this instance, it is not clear whether our data provide evidence for \( \mathcal{H}_0 \) (i.e., evidence of absence) or our data are simply not evidential for either hypothesis (i.e., absence of evidence). The Bayes factor gives us the ability to separate these two ideas. For example, if \( \text{BF}_{01} \) is large, we take this to represent evidence for the null -- that is, evidence for the absence of something.  Another benefit include the ability for the researcher to monitor evidence as the data accumulate. Though problematic for inference based on \( p \)-values, optional stopping does not affect the interpretation of Bayesian analyses [@rouder2014optional]. Thus, researchers may continuously monitor the "evidential flow" of their data and choose to stop whenever evidence is sufficiently compelling. Finally, Bayesian methods integrate testing with estimation, allowing the researcher to not only measure the evidence for one hypothesis over another, but to simultaneously obtain a probabilistic estimate of the effect size under the hypothesis that it is non-zero.

Given these benefits, what does a Bayesian hypothesis test require of the researcher? Like any hypothesis testing situation, the researcher must specify a null hypothesis \( \mathcal{H}_0 \) and an alternative hypothesis \( \mathcal{H}_1 \). In a Bayesian context, though, there are some specific considerations that govern how this choice is made. Essentially, the researcher must specify a *prior distribution* of effect sizes that are expected under \( \mathcal{H}_1 \). This allows \( \mathcal{H}_1 \) to makes predictions that can be evaluated against observed data. For example, the Bayesian \( t \)-test of @rouder2009bayesian uses a Cauchy distribution with scale \( r=1/\sqrt{2} \) as a default choice^[This prior satisfies several desirable properties for an objective, default prior; see the supplemental material for details.] for the prior distribution of effect sizes that are expected *if* there is a nonzero effect. Given this definition of \( \mathcal{H}_1 \) (i.e., as a distribution of expected effect sizes \( \delta \)), the null hypothesis \( \mathcal{H}_0 \) is specified as a specific point within this distribution, namely \( \delta=0 \). Thus, we can succinctly encode our hypotheses as:<!--
-->
\begin{align*}
\mathcal{H}_0: \delta &=0\\
\mathcal{H}_1: \delta &\sim \text{Cauchy}(1/\sqrt{2}).\\
\end{align*}
